Tailored Spiral In-Out Spectral-Spatial Water Suppression Pulses for Magnetic Resonance Spectroscopic Imaging

### 2020-08-14 Manuscript/Introduction.tex
edited wording

\par Multidimensional RF pulses have been widely used in many applications such as reduced field-of-view imaging \cite{rieseberg2002two}, transmit field inhomogeneity correction \cite{saekho2005small}, and susceptibility artifacts correction \cite{stenger2000three}. Parallel transmission (pTx) \cite{katscher2003transmit,zhu2004parallel} further strengthens the multidimensional pulses by providing extra spatial encoding power, shortening pulse duration, and reducing  specific absorption rate (SAR). However, current multidimensional pTx pulse design methods are based on a spatial domain formulation \cite{Grissom:2006:MRM,setsompop2008magnitude} that has prohibitive memory and computational requirements when the number of coils and the number of dimensions both become larger in current designs. 
Particularly, current spatial domain design methods map any given RF pulse $\mathbf{b}$ to its spatial excitation pattern $\mathbf{d}$ with a large system matrix $\mathbf{A}$, as:
\begin{equation*}
	\mathbf{d}=\mathbf{Ab}
\end{equation*}
The size of $\mathbf{A}$ is $N_s$ by $N_cN_t$, where $N_s$ is the number of spatial locations, $N_c$ is the number of coils, and $N_t$ is the number of RF samples. Since $N_s$ becomes unbearably large in designs such as 2D plus spectral and 3D problems, and $N_c$ increases with more advance hardware designs, this $\mathbf{A}$ matrix is typically too large to find the inverse in terms of memory. Therefore, such design problems need to be solved by iterative conjugate-gradient (CG) methods which avoid matrix inversion. However, even if no matrix inversions are performed, the large system matrix may still be memory-inefficient when used in many matrix multiplications over a large number of iterations, making the solving process very slow. Furthermore, the NUFFT-based pulse design methods can also be slow due to all the gridding steps. Although these can be eliminated in NUFFT-based image reconstruction using Toeplitz formulation, in RF pulse design the Toeplitz formulation does not exist since the inner product is over space. 
These slow design methods are especially undesirable when tailored pulses are needed to be designed for each subject during one's examination in the scanner. Typically, the subject-tailored pulse design has to be done within one minute in the pre-scan stage, so that the incorporation of subject-tailored pulses into currently used workflow is practical and desirable.

\par One application of subject-tailored multidimensional pTx pulses that is of particular interest of this dissertation is the 3D inner volume suppression (IVS) pulses for MR Corticography (MRCoG).MRCoG is a developing imaging technique which aims for submillimeter isotropic resolution whole-cortex and cortex-specific imaging. 
It is a promising technique to enable whole-brain functional and diffusion studies of the columnar and laminar subcortical structures, which are the fundamentals of higher-order brain functions. 
%The current columnar and laminar fMRI methods all use a small field of view in order to achieve sub-millimeter resolution. This does not allow studies to be performed across the whole cerebral cortex. On the other hand, current whole brain fMRI methods typically have spatial resolution greater than 1 mm, which is not sufficient for the studies of columnar and laminar structures. 
MRCoG will use IVS to enable highly accelerated imaging of the cortex, by reducing g-factor and suppressing physiological noise from ventricle CSF. Therefore, subject-tailored IVS pTx pulses are needed, which could be applied before each excitation and readout. Thanks to the high performance gradient system an the 24-channel transmit system of the developing MRCoG scanner, good IVS with 3D selective excitation pulses becomes feasible. However, this IVS pulse design problem is challenging due to its 3D nature and the large number of transmit channels.

>>>>

\par Multidimensional RF pulses have been widely used in many applications such as reduced field-of-view imaging \cite{rieseberg2002two}, transmit field inhomogeneity correction \cite{saekho2005small}, and susceptibility artifacts correction \cite{stenger2000three}. Parallel transmission (pTx) \cite{katscher2003transmit,zhu2004parallel} further strengthens the multidimensional pulses by providing extra spatial encoding power, shortening pulse duration, and reducing  specific absorption rate (SAR). However, current multidimensional pTx pulse design methods are based on a spatial domain formulation \cite{Grissom:2006:MRM,setsompop2008magnitude} that has prohibitive memory and computational requirements when the number of coils and the number of dimensions both become larger in current designs. 
Particularly, current spatial domain design methods map any given RF pulse $\mathbf{b}$ to its spatial excitation pattern $\mathbf{d}$ with a large system matrix $\mathbf{A}$, as:
\begin{equation*}
	\mathbf{d}=\mathbf{Ab}
\end{equation*}
The size of $\mathbf{A}$ is $N_s$ by $N_cN_t$, where $N_s$ is the number of spatial locations, $N_c$ is the number of coils, and $N_t$ is the number of RF samples. Since $N_s$ becomes unbearably large in designs such as 2D plus spectral and 3D problems, and $N_c$ increases with more advance hardware designs, 
This $\mathbf{A}$ matrix is typically too large to find the inverse in terms of memory. Therefore, such design problems need to be solved by iterative conjugate-gradient (CG) methods which avoid matrix inversion. However, even if no matrix inversions are performed, the large system matrix may still be memory-inefficient when used in many matrix multiplications over a large number of iterations, making the solving process very slow. Furthermore, the NUFFT-based pulse design methods can also be slow due to all the gridding steps. Although these can be eliminated in NUFFT-based image reconstruction using Toeplitz formulation, in RF pulse design the Toeplitz formulation does not exist since the inner product is over space. 
These slow design methods are especially undesirable when tailored pulses are needed to be designed for each subject during one's examination in the scanner. Typically, the subject-tailored pulse design has to be done within one minute in the pre-scan stage, so that the incorporation of fast design of subject-tailored pulses into currently used workflow is desirable.

\par One application of subject-tailored multidimensional pTx pulses that is of particular interest of this paper is the 3D inner volume suppression (IVS) pulses for MR Corticography (MRCoG).MRCoG is a developing imaging technique which aims for submillimeter isotropic resolution whole-cortex and cortex-specific imaging. 
It is a promising technique to enable whole-brain functional and diffusion studies of the columnar and laminar subcortical structures, which are the fundamentals of higher-order brain functions. 
%The current columnar and laminar fMRI methods all use a small field of view in order to achieve sub-millimeter resolution. This does not allow studies to be performed across the whole cerebral cortex. On the other hand, current whole brain fMRI methods typically have spatial resolution greater than 1 mm, which is not sufficient for the studies of columnar and laminar structures. 
MRCoG will use IVS to enable highly accelerated imaging of the cortex, by reducing g-factor and suppressing physiological noise from ventricle CSF. Therefore, subject-tailored IVS pTx pulses are needed, which could be applied before each excitation and readout. Thanks to the high performance gradient system an the 24-channel transmit system of the developing MRCoG scanner, good IVS with 3D selective excitation pulses becomes feasible. However, this IVS pulse design problem is challenging due to its 3D nature and the large number of transmit channels.

### 2020-08-14 Manuscript/Theory.tex
edited wording

\begin{equation}
\mathbf{S}^{H}\mathbf{S} = 
\begin{pmatrix}
\mathbf{S}_1^{H}\mathbf{S}_1 & \cdots & \mathbf{S}_1^{H}\mathbf{S}_{N_c} \\
\vdots  &  \ddots & \vdots  \\
\mathbf{S}_{N_c}^{H}\mathbf{S}_1 & \cdots & \mathbf{S}_{N_c}^{H}\mathbf{S}_{N_c} 
\end{pmatrix}
\end{equation}

@@

\subsection*{ Off-Resonance correction}
\par As shown in Equation \ref{eq:SpatialDomainEquation} when we introduced spatial domain design in Chapter \ref{chap:SpatialDesign}, the equation connecting a pTx pulse to its excitation pattern has an off-resonance dependent term 
$e^{\imath \gamma \Delta B_0(\mathbf{x})(t-T)}$. In order to model this off-resonance dependent term in our k-space domain design, we adapted a fast implementation of the off-resonance compensation with time segmentation approximation and weighted least-squares interpolators \cite{fessler2005toeplitz}. 
By segmenting the RF pulse into multiple segments, this method disentangles the space-and-time dependent phase accrual induced by main field inhomogeneity into either space- or time- dependent terms, and therefore a same system matrix can be applied within each segment. Mathematically, it is
\begin{equation*}
e^{\imath \Delta\omega_j (t_i-T) }=\sum_{l=1}^{L} b_{l}(t_i) e^{\imath \theta_{l}(\Delta\omega_j)}
\end{equation*}
where $i=1\dots Nt$ indexes the RF samples, $j=1\dots Ns$ indexes the spatial locations, and $l$ indexes the time segments. The left term is the problematic space-and-time dependent term, with $\Delta\omega_j$ representing the off-resonance at the location-$j$ obtained from a field map and $T$ being the pulse duration. $b_{il}=b_{l}(t_i)$ are the time-dependent-only weights called temporal interpolators applied upon RF samples. $c_{lj}=e^{\imath \theta_{l}(\Delta\omega_j)}$ are the space-dependent-only phase accrual called spatial interpolators, applied upon the excitation pattern of each weighted RF segment. 
Therefore, the off-resonance correcting RF can be obtained by 
\begin{equation*}
\mathbf{b}=\sum_{l=1}^{L}diag\{\mathbf{b}_l\}\mathbf{W}\mathcal{F}(diag\{\mathbf{c}_l\}\mathbf{d})
\end{equation*}
where $\mathbf{b}_l$ and $\mathbf{c}_l$ are the vectorized interpoloators for the $l$-th segment.

>>>>

\begin{equation}\label{eq:SHS_blocks}
\mathbf{S}^{H}\mathbf{S} = 
\begin{pmatrix}
\mathbf{S}_1^{H}\mathbf{S}_1 & \cdots & \mathbf{S}_1^{H}\mathbf{S}_{N_c} \\
\vdots  &  \ddots & \vdots  \\
\mathbf{S}_{N_c}^{H}\mathbf{S}_1 & \cdots & \mathbf{S}_{N_c}^{H}\mathbf{S}_{N_c} 
\end{pmatrix}
\end{equation}

@@

\subsection*{ Off-Resonance correction}
%\par As shown in Equation \ref{eq:SpatialDomainEquation} when we introduced spatial domain design in Chapter \ref{chap:SpatialDesign}, the equation connecting a pTx pulse to its excitation pattern has an off-resonance dependent term 
$e^{\imath \gamma \Delta B_0(\mathbf{x})(t-T)}$. 
In order to model the off-resonance into our k-space domain design, we adapted a fast implementation of the off-resonance compensation with time segmentation approximation and weighted least-squares interpolators \cite{fessler2005toeplitz}. 
By segmenting the RF pulse into multiple segments, this method disentangles the space-and-time dependent phase accrual induced by main field inhomogeneity into either space- or time- dependent terms, and therefore a same system matrix can be applied within each segment. Mathematically, it is
\begin{equation*}
e^{\imath \Delta\omega_j (t_i-T) }\approx\sum_{l=1}^{L} g_{l}(t_i) e^{\imath \theta_{l}(\Delta\omega_j)}
\end{equation*}
where $i=1\dots Nt$ indexes the RF samples, $j=1\dots Ns$ indexes the spatial locations, and $l$ indexes the time segments. The left term is the problematic space-and-time dependent term, with $\Delta\omega_j$ representing the off-resonance at the location-$j$ obtained from a field map and $T$ being the pulse duration. $g_{il}=b_{l}(t_i)$ are the time-dependent-only weights called temporal interpolators applied upon RF samples. $h_{lj}=e^{\imath \theta_{l}(\Delta\omega_j)}$ are the space-dependent-only phase accrual called spatial interpolators, applied upon the excitation pattern of each weighted RF segment. Therefore, the spatial domain formulation becomes
\begin{equation*}
\mathbf{d}=\sum_{l=1}^{L}diag\{\mathbf{h}_l\}\mathbf{A}(diag\{\mathbf{g}_l\}\mathbf{b})
\end{equation*}
The spatial interpolators can be absorbed into the sensitivity maps. The $\mathbf{S}$ matrix in Equation \ref{eq:MatrixForm} will then become 
\begin{equation*}
\mathbf{S_{offres}}=\sum_{l=1}^{L}\mathbf{S}_l diag\{\mathbf{g}_l\}
\end{equation*}
where the columns of the $\mathbf{S}_l$ matrix, instead of being the aforementioned k-space sensitivity maps, are now the Fourier transform of the dot products of the spatial $B_1^+$ maps and the spatial interpolatros of the $l$-th segment, shifted to corresponding excitation k-space locations. 
The efficient construction of $\mathbf{S_{offres}}^{H}\mathbf{S_{offres}}$ follows the same concept mentioned before. 
But with off-resonance, for the elements within an element matrix $\mathbf{S}_i^{H}\mathbf{S}_j$ in Equation \ref{eq:SHS_blocks}, instead of being interpolated from one densely sampled Fourier transform map, which is of the dot product of the spatial domain $B_1^+$ maps of the $i$-th and $j$-th coils, now they are interpolated from $L(L+1)/2$ different Fourier transform maps, each of which has underlying spatial $B_1^+$ maps with two sets of spatial interpolators absorbed into them. For the $(m,n)$-th element within the element matrix $\mathbf{S}_i^{H}\mathbf{S}_j$, the $L(L+1)/2$ interpolated values were then summed with weights $g_{l_i}(t_m)g_{l_j}(t_n)$, where $l_i$ and $l_j$ are the two segment numbers of the spatial interpolators observed into the corresponding $B_1^+$ maps. 
Compared to the efficient construction method without off-resonance, taking off-resonance into account increases the total number of FFT operations form $N_c(N_c+1)/2$ to $N_cL(N_cL+1)/2$, and increased the number of interpolations needed for each element from 1 to $L(L+1)/2$.
